### 2021-03-29 outputs/paper/paper.Rmd
repo setup

---
title: "Climate Perceptions in Toronto"
author: "Lorena Almaraz De La Garza"
date: "2021 04 19"
output:
  bookdown::pdf_document2: default
subtitle: "with data from Open Data Toronto"
toc: FALSE
thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions] (https://github.com/lalmaraz/toronto_climate_perceptions).'
bibliography: references.bib
abstract: X
---

# Introduction

```{r libs, include = FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load libraries
library(bookdown)
library(ggmap)
library(gt)
library(here)
library(httr)
library(janitor)
library(kableExtra)
library(knitr)
library(opendatatoronto)
library(readxl)
packageVersion("readxl")
library(rio)
library(tidyverse)

# Find and get data

questions <- search_packages("climate") %>% 
  list_package_resources() %>% 
  filter(grepl("Readme Data Field Names", name)) %>% 
  get_resource()

raw_data <- search_packages("climate") %>% 
  list_package_resources() %>% 
  filter(grepl("City of Toronto Climate Perceptions SPSS v1", name)) %>% 
  get_resource()

# Save data
write_csv(raw_data,"inputs/data/raw_data.csv")

# Read in data
all_data <- read_csv("inputs/data/raw_data.csv")
```

# Data

# Discussion

## Limitations

## Next Steps

\newpage
# Appendix 1 {#appendix1}
```

\newpage
# References

### 2021-04-04 outputs/paper/paper.Rmd
incl paper sections

@@
 author: "Lorena Almaraz De La Garza"
-date: "2021 04 19"
+date: "April 19th, 2021"
+keywords: "X, Y, Z"
 output:
@@
 toc: FALSE
-thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions] (https://github.com/lalmaraz/toronto_climate_perceptions).'
+thanks: 'Code and data are available in this GitHub repository: [lalmaraz/toronto_climate_perceptions](https://github.com/lalmaraz/toronto_climate_perceptions).'
 bibliography: references.bib
-abstract: X
+abstract: "The abstract text occupies this beautiful space right here. \\par
+ \\textbf{Keywords:} search, stackoverflow, find, workaround"
+
 ---
-
-# Introduction
 
@@
 library(readxl)
-packageVersion("readxl")
 library(rio)
 library(tidyverse)
+```
 
-# Find and get data
-
-questions <- search_packages("climate") %>% 
-  list_package_resources() %>% 
-  filter(grepl("Readme Data Field Names", name)) %>% 
-  get_resource()
-
-raw_data <- search_packages("climate") %>% 
-  list_package_resources() %>% 
-  filter(grepl("City of Toronto Climate Perceptions SPSS v1", name)) %>% 
-  get_resource()
-
-# Save data
-write_csv(raw_data,"inputs/data/raw_data.csv")
-
-# Read in data
-all_data <- read_csv("inputs/data/raw_data.csv")
-```
+# Introduction
+[@R]
 
@@
 
+## Methodology
+## Results
+
 # Discussion
 
+## Findings
+## Implications
 ## Limitations
+## Future Work
 
-## Next Steps
+```{r echo=FALSE, message=FALSE, warning=TRUE}
+
+```
 
@@
 # Appendix 1 {#appendix1}
-```
 

### 2021-04-07 outputs/paper/paper.Rmd
intro draft 1

@@
 ---
-title: "Climate Perceptions in Toronto"
+title: "Climate Perceptions in Toronto: [The effect of age on self-reported level of concern]"
 author: "Lorena Almaraz De La Garza"
@@
 bibliography: references.bib
-abstract: "The abstract text occupies this beautiful space right here. \\par
- \\textbf{Keywords:} search, stackoverflow, find, workaround"
+abstract: "This report presents a secondary analysis on the City of Toronto's publicly-accessible data on residents' perceptions about climate change.
+The results indicate there is a [negative/positive] association between respondents' age and self-reported level of concern of climate change, and a [negative/positive] association between respondents' gender and self-reported level of concern of climate change.
+The analysis consists of logistic regression conducted with the statistical programming language R.
+Effectively addressing the ongoing threat of climate change demands collaboration and data-informed decision-making.
+[The results] contribute to our understanding of the issue at a local level, enabling effective action.
+ \\par
+ \\textbf{Keywords:} climate change, climate perceptions, climate concerns, climate action, age and climate, youth and climate, toronto residents, open data toronto"
 
@@
 # Introduction
-[@R]
+
+Although scientific research on the impacts of climate change has been conducted for decades, the last few years have seen a surge of public participation and activism.
+In particular, youth climate leaders, with unwavering motivation and forthright campaigns, have guided and inspired thousands around the world [@unicef].
+In their demands they often ask that adults enact intergenerational equity - the principle that states that every generation, present and future, "holds the Earth in common" [@oxford] - and wake up from their profit-driven ignorance and comfort [@villasenor].
+But how does age really affect self-reported concern of climate change?
+To answer this question, the following pages present a secondary analysis on survey data collected by Environics, commissioned by the City of Toronto, on residents' perceptions on climate change.
+Using the statistical programming language R, this study performs logistic regression to determine whether a correlation exists between two key variables: age and self-reported level of concern about climate change.
+Informed by the youth climate movement, the hypothesis is that younger generations are more inclined to self-describe as concerned about climate change.
+
+Drawing upon the activism that advocates for an intersectional approach to climate action, a similar regression analysis is also performed on the gender variable.
+Intersectionality, as defined by Kimberle Crenshaw and described by Nina Lykke, is the idea that discrimination and systemic inequalities overlap – or intersect – across factors like gender, race, and class, and do not exist in isolation [@lykke].
+Intersectional environmentalism, therefore, promotes climate action that recognizes the ways in which climate change disproportionately affects the already disadvantaged [@madson].
+In this case, the hypothesis is that participants that identified as female will have a greater likelihood of reporting a greater level of concern of climate change.
+
+The results provide valuable insight.
+[Results will be here.] As argued by Jackson and De Coninck, effective climate action, particularly in adaptation and resilience, depends on active participation at a local level [@jackson].
+With this in mind, this paper advocates for the use of open data practices by local governments, including municipalities, to promote data-informed decision-making.
+For civic engagement and grassroots organizing, access to reliable data is vital.
+It allows organizers to direct outreach efforts and to allocate precious resources efficiently.
+Due to the scale and complexity of the challenges posed by climate change, ensuring the public has access to quality data can prove immensely beneficial.
+As beautifully illustrated by Kieran Healy, the crowdsourced data and data analyses that continue to uncover the effects of COVID-19 demonstrate the power of collaboration and open data in addressing challenges of unforeseen scales [@healy].
+Similarly, climate change - and effective climate action - have a lot to gain from public ingenuity fueled by open data.
+
+The paper is organized as follows.
+The Data section presents an overview of the original dataset and the variables relevant to this study.
+It also provides critical commentary on the limitations of the original survey design.
+Finally, it compares the surveyed sample to the provincial and national population.
+The Methodology section discusses the strengths and weaknesses of the implemented model, as well as of the models that were considered as alternatives.
+It includes details on how the data were re-coded to suit the model and how non-response is addressed.
+The Results section explains the baseline category in the model and the coefficients from each analysis.
+Finally, the Discussion section is comprised of four subsections: Findings, Ethics & Implications, Limitations, and Future Work, situating the study within a broader context.
 
 # Data
+- R [@R] & packages
+- Origin: Environics commissioned by the City of Toronto
+- Sample:
+- Variables: level of concern [yes/no], level of information [yes/no], age [four groups from report], neighbourhood [wards], gender [m/f], dwellings, own/rent, income level, education level
+- Survey limitations/biases: non-response bias (unit/item)
+- Is data missing in patterns?
+- Respondents' demographics compared to broader (provincial/national) population
+- Nice graphs and tables that show the data as I found it
 
 ## Methodology
+- Logistic regression to show correlation
+- This is a good choice because [strengths] 
+- Clear equation
+- Applied to these data specifically, explain notation
+- The things this *does not* tell us [weaknesses]
+- Even accounting for these ^, this is a reasonable approach because [reasons/alternative models' limitations]
+- How non-response is dealt with here
+- X variables re-coded into binaries
+
 ## Results
+- Presented with nothing but the best, A Nice Table
+- Baseline category [refactor if needed]
+- Coefficients [exponentiated]
+- Age will be in groups (so +1 age means +1 age GROUP)
+- For same conditions [education/income/dwellings/etc], change of 1 in a variable[age/gender] is X% more/less likely to self-report more/less concern? More/less informed?
+- What else was in the data [refer to findings from EDA script]? Patterns/trends? Any unasked but answered questions?
 
@@
 ## Findings
-## Implications
+- [Finding 1]
+- [Finding 2]
+- [Finding 3]
+
+## Ethics & Implications
+- Ethics
+- What exactly did the city do with these data?
+[Women4Climate programme, Wild About Bees, Neighbourhood Climate Action Champions... within TransformTO/Live Green Toronto/Env Grants & Incentives]
+- Open data & civic organizing (1240W10P1)
+- Value of local data-informed action for climate change [may or may not self-promote here, prob won't. but maybe?]
+
 ## Limitations
+- Observational data & general limitations of any model, THERE IS NO TRUTH!
+- Speaking on causality [and why I can't/won't]
+- Several things potentially at play here not studied or even acknowledged: research on political affiliation and climate perception, etc
+- Berkson's when looking just at Toronto?
+- Quick look at the city's report (how do they draw conclusions from such a limited dataset?)
+
 ## Future Work
+- MRP for Ontario-wide perceptions?
 

### 2021-04-27 scripts/EDA.R
Paper and scripts update

@@ -11,6 +11,8 @@ library(opendatatoronto)
 library(RCurl)
 library(here)
 library(tidyverse)
+library(performance)
+
 
 # Based on Michael Chong's excellent work, as published in Telling Stories With Data (Alexander, 2021) here:
 # https://www.tellingstorieswithdata.com/exploratory-data-analysis.html#case-study---opinions-about-a-casino-in-toronto
@@ -170,3 +172,15 @@ summary(boomer_concern_global_glm)
 # Informed
 boomer_informed_glm <- glm(informed_binary ~ boomer_status + gender + education_group + income_group, data = climate_perceptions_clean, family = "binomial")
 summary(boomer_informed_glm)
+
+
+# Assess model fit
+r2_mcfadden(boomer_informed_glm)
+# takes deviance from intercept-only (no predictor) and subtracts from full model
+# creates index of proportionate increase in the model fit from including predictors
+
+r2_coxsnell(m_concern_global_glm)
+
+r2_nagelkerke(m_concern_global_glm)
+
+r2_tjur(m_concern_global_glm)

### 2021-04-27 scripts/compare_demographics.R

@@ -12,6 +12,7 @@ library(tidyverse)
 library(RCurl)
 library(janitor)
 
+
 # Get data
 census_2016 <- getURL("https://www12.statcan.gc.ca/census-recensement/2016/dp-pd/prof/details/download-telecharger/current-actuelle.cfm?Lang=E&Geo1=CSD&Code1=3520005&Geo2=PR&Code2=35&B1=All&type=0&FILETYPE=CSV")
 
@@ -54,7 +55,7 @@ age_data <- age_data %>%
   na.omit() %>% 
   slice(1:(n()-1)) # Remove irrelevant row
 
-# Make city/province table
+# Make city/province demographics table
 totals_table_data <- age_data %>%
   select(4, 5, 7, 8, 9) %>%
   group_by(age_group) %>%
@@ -63,6 +64,171 @@ totals_table_data <- age_data %>%
             male_on = sum(male_on),
             female_on = sum(female_on))
 
-city_province_totals <- census_2016_data[8,]
+# Make sample demographics table
+climate_perceptions <- read.csv("outputs/data/climate_data.csv")
 
-# TO DO: make graphs w percentages/proportion of male / female per age group and compare to study sample
+# Add column for age groups as determined in original report
+climate_perceptions <- 
+  climate_perceptions %>%
+  mutate(age_group = case_when(
+    age >= 18 & age <= 34 ~ 0,
+    age >= 35 & age <= 49 ~ 1,
+    age >= 50 & age <= 64 ~ 2,
+    age >= 65 ~ 3
+  ))
+
+# Vector with male counts per age group from survey
+ms <- climate_perceptions %>% 
+  select (gender, age, age_group) %>% 
+  group_by(age_group) %>% 
+  filter(gender == "Man") %>% summarize(male_s = n())
+
+# Vector with female counts per age group from survey
+fs <- climate_perceptions %>% 
+  select (gender, age, age_group) %>% 
+  group_by(age_group) %>% 
+  filter(gender == "Woman") %>% summarize(female_s = n())
+
+# Join
+ms_fs <- inner_join(ms, fs, by = "age_group")
+
+totals_table_data <- inner_join(ms_fs, totals_table_data, by = "age_group") %>% 
+  mutate(total_s = 0,
+         total_to = 0,
+         total_on = 0)
+
+# New columns and rows for total counts (not separated by gender)
+totals_table_data$total_s=rowSums(cbind(totals_table_data$male_s,totals_table_data$female_s))
[…]

### 2021-04-27 scripts/map_regions.R

@@ -12,7 +12,10 @@ library(here)
 library(rvest)
 library(RCurl)
 library(tidyverse)
-library(xml2)
+library(xml2) 
+library(ggmap)
+library(osmdata)
+library(sf)
 
 # Get data
 postal_codes <- read_html("https://www.zipcodesonline.com/2020/06/postal-code-of-toronto-in-2020.html")
@@ -26,8 +29,6 @@ postal_codes <- postal_codes %>%
   html_text()
 
 website_text <- tibble(postal_codes = postal_codes)
-
-head(postal_codes)
 
 # Slice only relevant info
 website_text <- website_text %>% 
@@ -65,9 +66,6 @@ postal_codes_clean <- postal_codes_clean %>%
   mutate(postal_code = str_sub(postal_code,0,3)) %>%   # Keep only first postal code if multiple
   unique()
 
-# Reminder: Run EDA script first, it generates the file read here:
-climate_perceptions_clean <- read.csv("outputs/data/climate_data_clean.csv")
-
 # Get latitude and longitude data
 # CA_ON_full.txt is an excerpt of CA_full, available as .zip here: https://download.geonames.org/export/zip/
 # CC GeoNames (www.geonames.org)
@@ -81,7 +79,7 @@ coordinates_to <- coordinates_on %>%
   group_by(V2) %>% 
   summarise(V2 = V2,
             V10 = ave(V10),
-            V11 = ave(V11)) %>% 
+            V11 = ave(V11)) %>% # Don't need exact location, an average is fine
   unique() %>% 
   rename(
     postal_code = V2,
@@ -89,9 +87,47 @@ coordinates_to <- coordinates_on %>%
     longitude = V11
   )
 
-# TO DO: Check postal codes in climate_perceptions_clean against postal_codes_clean,
-# make column in 1st that assigns the appropriate region
+# Join dataframes and add missing postal codes and coordinates manually
+postal_codes_coordinates <- inner_join(postal_codes_clean, coordinates_to, by = "postal_code") %>% 
+  add_row(postal_code = "M2N", region = "North York", latitude = 43.77093629736116, longitude = -79.4132495121774) %>% 
+  add_row(postal_code = "M2R", region = "North York", latitude = 43.77913556475642, longitude = -79.4442928405029) %>% 
+  add_row(postal_code = "M3C", region = "North York", latitude = 43.722655855005605, longitude = -79.34073514475693) %>% 
[…]

### 2021-04-28 shiny/app.R
shiny app

@@ -1,0 +1,101 @@
+
+# library(shiny)
+library(ggmap)
+library(osmdata)
+library(RCurl)
+# library(tidyverse)
+# library(rsconnect)
+
+# load data
+map_text <- getURL("https://raw.githubusercontent.com/lalmaraz/toronto_climate_perceptions/main/outputs/data/map_data_full.csv")
+map_data <- read.csv(text = map_text)
+map_tiles <- get_map(getbb("Toronto"),maptype = "terrain")
+
+map_data[map_data == 0] <- NA
+
+ui <- fluidPage(
+    titlePanel("Climate Perceptions Study: Demographics"),
+    sidebarLayout(
+        sidebarPanel("Select an item from the drop down.",
+                     selectInput("dropdown", label = "Characteristics", choices = list("AGE" = list("18-34", "35-49", "50-64", "65+"),
+                                                                                       "GENDER" = list("Male", "Female"),
+                                                                                       "EDUCATION" = list("High school or less", 
+                                                                                                          "Some community college", 
+                                                                                                          "Completed community college",
+                                                                                                          "Some university",
+                                                                                                          "Completed undergraduate degree",
+                                                                                                          "Post graduate/professional school"),
+                                                                                       "INCOME" = list("Under $40,000",
+                                                                                                       "$40,001 to $60,000",
+                                                                                                       "$60,001 to $80,000",
+                                                                                                       "$80,001 to $100,000",
+                                                                                                       "$100,001 to $150,000",
+                                                                                                       "More than $150,000"),
+                                                                                       "CONCERNED ABOUT CLIMATE CHANGE" = list("Local Impacts: Yes", "Local Impacts: No", "Global Impacts: Yes", "Global Impacts: No"),
[…]
